Make any stupid plot you want

### 2021-01-13 draft/sections/math.tex
reworked\reorganized example

In this section we introduce a mathematical description of the visualization pipeline where artist $\mathfrak{A}$ functions transform data space $\mathfrak{E}$ to an intermediate representation in a prerendered display space $\mathfrak{H}$.

\begin{equation}
    \label{eq:artist}
    \mathcal{A}: \mathcal{E} \rightarrow \mathcal{H}
\end{equation}

We use fiber bundles\cite{FiberBundle2020, rowlandFiberBundle} to model data and graphics because they allow to seperate concerns

%%find word to put here topological total space that encodes points and a topological base space that can encode the connectivity of those points:
\begin{itemize}
\item $E$ is a locally trivial fiber bundle over $K$ representing data space.
\item $H$ is a fiber bundle over $S$ representing visual space
\item $K$ and $S$ are a triangulizable topological space or a CW complex encoding the connectivity of points in $E$ and $H$ respectively
\end{itemize}

@@

\subsection{Data Space $E$}
As proposed by Butler \cite{butlerVectorBundleClassesForm1992,butlerVisualizationModelBased1989}, we model data as a fiber bundle ($E$, $K$, $\pi$, $F$)

@@

\paragraph{Example}

\begin{figure}[H]
    \label{fig:data_fiber_bundle}
    \includegraphics[width=1\linewidth]{figures/sections/math/fiberbundle.png}
    \caption{$F$ is an $n\times m$ space of points. The section $\tau_1$ returns the blue points,  while $\tau_2$ returns the purple points. $\Gamma(E)$ is the set of all sections, including $\tau_1$ and $\tau_2$}.  
\end{figure}

The fiber bundle in figure~\ref{fig:data_fiber_bundle} can encode different types of data. $F$ is a cartesian product of the codomains of 2 variables and $K$ is a line on the interval [0,1]. The sections $\tau_{1}$ and $\tau_{2}$ are different sets of data values in the space.

\begin{description}
    \item[distribution] $F$ is the cartesian cross product of temperature values and the range of probabilities [0,1] such that every point in $E$ is a (temperature, probabilty) pair. $\tau_1$ and $\tau_2$ each return a different distribution of temperatures. 
    \item[timeseries]  $F = \{\text{temperature}\} \times \{\text{timestamps}\}$, such that every point is a (temperature, time) pair. $\tau_1$ and $\tau_2$ each return different timeseries of temperature.
\end{description}

\subsubsection{Base Space $\mathcal{K}$}
\label{sec:base_data}
\begin{figure}[H]
    %%\includegraphics[\textwidth]{}
    \label{fig:base_space_types}
    %%% pull up data structures like discrete, continous, etc w/ base space 
    %%% little drawing dots, line, plane, cube
    \caption{The topological base space $K$ encodes the connectivity of the data space, for example if the data is independent points or a map or on a sphere}
\end{figure}

Data can be discrete observations, timeseries, maps, fields \cite{munznerWhatDataAbstraction2014} and $K$ is a set of points $k$ that can act as keys from a representative space, such as seen in figure~\ref{fig:base_space_types}, to the data values in $E$ \cite{munznerWhatDataAbstraction2014}.

\begin{figure}[H]
    \label{fig:base_space_div}
    \includegraphics[width=.5\linewidth]{figures/sections/math/k_qspace.png}
    %%include second figure where fiber goes otherway
\end{figure}

$K$ and $F$ are not intrinsic to $E$, rather they are choices in how $E$ is subdivided\cite{QuotientSpaceTopology2020,QuotientSpaceTopology2020}. In figure~\ref{fig:base_space_div} we can divide a rectangular base space such that there is a short fiber and long base space or a long fiber and short base space. This is analogous to long and wide forms of the same table \cite{wickham2014tidy}.

\paragraph{Example}
\begin{figure}[ht]
    \includegraphics[width=\textwidth]{figures/sections/math/temp_1k.png}
    %% add box around neighboring P and Map
    \label{fig:k_data}
\end{figure}

\begin{figure}[H]
    \includegraphics[width=\textwidth]{figures/sections/math/temp_2k.png}
\end{figure}

in figure~\ref{fig:k_data}, temperature is the only one data field in $r$ but the $K$ base spaces are different. subfig[1] is a timeseries, so the temperature in $r$ at time $t$ is dependent on the temperature in $r_{t-1}$ and the temperature in $r_{t+1}$ is dependent on  $r_t$; this connectivity is expressed as a one dimensional $K$ where $K$ is the number line. In the case of the map, every temperature in $r$ is dependent on its nearest neighbors on the plane, and one way to express this is by encoding $K$ as a plane. $K$ does not know the time or latitude or longitude of the point as those are metadata variables describing the $k$ rather than the value of $k$. The mapping $\tau: K \rightarrow E$ provides the binding between the key $k \in K$ and the value $r$ in $E$ \cite{munznerChDataAbstraction}.

@@

\subsubsection{Example}
\begin{figure}[H]
    \includegraphics[width=0.2\linewidth]{figures/sections/math/temp_2f.png}
    \label{fig:}
\end{figure}
\begin{figure}[H]
    \includegraphics[width=0.2\linewidth]{figures/sections/math/temp_3f.png}
\end{figure}

%% pull out measurement
The fiber in figure~\ref{fig:temp} is the space of possible temperature values in degrees celsius, such that $F=[temp_{min}, temp_{max}]$ and is named \textrm{Temp}. In figure~\ref{fig:temp_time} \textrm{time} is encoded as a second dimension. This means that the set of possible values $F$ with $C=\{\textrm{Temp}, \textrm{Time}\}$:

\begin{equation}
F = [temp_{min}, temp_{max}] \times [time_{min}, time_{max}]
\end{equation}

and the function $\tau$ that retrieves records from $F$ is

\begin{align}
\tau(k) =(k, (r: \textrm{Temp}\rightarrow &temp, r: \textrm{Time}\rightarrow time))\\
&temp \in [temp_{min}, temp_{max}], time \in [time_{min}, time_{max}])
\end{align}

Since $\tau(k)=(k, r)$, $temp$ is bound to a named data field and $sigma$ binds $temp$ to a temperature data type.

\subsubsection{Sheaf and Stalk}
%% add motivation - zoom, subsection, streaming, also for realizing visualization, restricting to that piece - need to know near point w/o

Often a graphic may need to be updated with live data or support zooming in on a segement of the dataset; to support working with a subset of data, we can use the sheaf $\mathcal{O}(E)$:

\begin{equation}
    \label{eq:sheaf}
    \begin{tikzcd}
        \iota^*E \arrow[d, "\pi"']           & E \arrow[d, "\pi"'] \arrow[l, "\iota^*"']         \\
        U \arrow[u, "\iota^*\tau"', bend right] & K \arrow[u, "\tau"', bend right] \arrow[l, "\iota"']
        \end{tikzcd}
\end{equation}

As shown in equation~\ref{eq:fiber_local_trivial}, there is a local space $U \subset K$ around every $k$. The inclusion map $\iota: U \rightarrow K$ is pulled back such that $\iota^{*}E$ is the space of $E$ restricted over $K$. The localized section of fibers $\iota^*\tau: U \rightarrow \iota^*E$ is the sheaf with a germ of $\xi^*\tau$. The neighborhood of points $k_i$ surrounding the point $k$ the sheaf lies over is the stalk $\mathcal{F}_b$ \cite{StalkSheaf2019,spanier1989algebraic}.

@@

@@

$M = M_{1} \times M_{2} \times \ldots \times M_{n}$
color X xpos (Cross product of monoid actions over V/E), cartesian product b/c independent. $M$ over $E$ was translated to acting over $\Gamma(V)$ of these act on $\Gamma(V)$
$\nu$ translates the sets, constraint on $\nu$ is equivariance of M. Is limited to the $Q(\Gamma(V))\subset of \Gamma(H)$ because not all visual tramsforms ($\mu$) are supported by all  $\rho$.
%%insert little thickness sketch maybe?O->(O)
So we define the actions $M$ on the image of $Q$, $Q(\Gamma(V)) = Y$. We can backward define our actions?

>>>>

In this section we introduce a mathematical description of the visualization pipeline where artist $\mathscr{A}$ functions transform data space $\mathscr{E}$ to an intermediate representation in a prerendered graphic space $\mathscr{H}$.

\begin{equation}
    \label{eq:artist}
    \mathscr{A}: \mathscr{E} \rightarrow \mathscr{H}
\end{equation}

We use fiber bundles\cite{FiberBundle2020, rowlandFiberBundle} to model data and graphics because they allow us to seperate the fields in a dataset from how the values in those fields are connected to each other:
\begin{itemize}
\item $E$ is a locally trivial fiber bundle over $K$ representing data space.
\item $H$ is a fiber bundle over $S$ representing visual space
\item $K$ and $S$ are a triangulizable topological space or a CW complex encoding the connectivity of points in $E$ and $H$ respectively
\end{itemize}

@@

\subsection{Data Space}
As proposed by Butler \cite{butlerVectorBundleClassesForm1992,butlerVisualizationModelBased1989}, we model data as a fiber bundle ($E$, $K$, $\pi$, $F$)

@@

\subsubsection{Base Space $\mathcal{K}$}
\label{sec:base_data}
\begin{figure}[H]
    \includegraphics[width=\textwidth]{figures/sections/math/k_different_types.png}
    \label{fig:base_space_types}
    \caption{The topological base space $K$ encodes the connectivity of the data space, for example if the data is independent points or a map or on a sphere}
\end{figure}

Datasets have a semantic topology such as the values are interpreted as discrete observations or part of a timeseries or map, or nodes in a networks \cite{munznerWhatDataAbstraction2014,geveci2012vtk}. As illustrated in fig~\ref{fig:base_space_div}, $K$ is this underlying structure. $K$ does not directly know the values; instead it is the sections that define the lookup between keys $k \in K$ and the corresponding values in $E$.

\begin{figure}[H]
    \label{fig:base_space_div}
    \includegraphics[width=.5\linewidth]{figures/sections/math/k_qspace.png}
    %%include second figure where fiber goes otherway
\end{figure}
The topology $K$ and the fields $F$ are determined by how $E$ is subdivided\cite{QuotientSpaceTopology2020,QuotientSpaceTopology2020}. In figure~\ref{fig:base_space_div} we can divide a rectangular base space such that there is a short fiber and long base space or a long fiber and short base space. This is analogous to long and wide forms of the same table \cite{wickham2014tidy}.

@@

\subsubsection{Sheaf and Stalk}

Often a graphic may need to be updated with live data or support zooming in on a segement of the dataset; to support working with a subset of data, we can use the sheaf $\mathcal{O}(E)$

\begin{equation}
    \label{eq:sheaf}
    \begin{tikzcd}
        \iota^*E \arrow[d, "\pi"']           & E \arrow[d, "\pi"'] \arrow[l, "\iota^*"']         \\
        U \arrow[u, "\iota^*\tau"', bend right] & K \arrow[u, "\tau"', bend right] \arrow[l, "\iota"']
        \end{tikzcd}
\end{equation}
which is the localized section of fibers $\iota^*\tau: U \rightarrow \iota^*E$ pulled back via the inclusion map $\iota: U \rightarrow K$. The localized section is the germ $\xi^*\tau$. The neighborhood of points $k_i$ surrounding the point $k$ the sheaf lies over is the stalk $\mathcal{F}_b$ \cite{StalkSheaf2019,spanier1989algebraic}.

@@

\subsubsection{Example: Temperature}
\textcolor{red}{Moved \& walked through b/c was getting clunky to not have terms yet}\\
The fiber bundle model is flexible enough to express some of the many different forms that temperature data can come in.

\begin{figure}[H]
    \begin{subfigure}{.5\textwidth}
        \includegraphics[width=\textwidth]{figures/sections/math/temp_1k.png}
        %% add box around neighboring P and Map
        \label{fig:base_example_line}
    \end{subfigure}
    \begin{subfigure}{.5\textwidth}
        \includegraphics[width=\textwidth]{figures/sections/math/temp_2k.png}
        \label{fig:base_example_plane}
    \end{subfigure}
    \label{fig:base_example}
    \caption{These two datasets have the same fiber of temperature but different base spaces. In figure~\ref{fig:base_example_line} the temperature values are 1D continuous, while in figure~\ref{fig:base_example_plane} the temperature values are 2D continuous.}
\end{figure}

The datasets in figure~\ref{fig:base_example} have identical fibers that encode a set of temperature values. In figure~\ref{fig:base_example_line} the temperatures lie on a line such that a section could return a timeseries or a distribution. In figure~\ref{fig:base_example_plane}, the temperatures lie on a 2D continuous plane; a section could return a map or contour. Because the fiber is 1D, it does not encode metadata

\begin{figure}[H]
    \begin{subfigure}{.5\textwidth}
        \includegraphics[width=\textwidth]{figures/sections/math/temp_2f.png}
        \label{fig:fiber_example_plane}
    \end{subfigure}
    \begin{subfigure}{.5\textwidth}
        \includegraphics[width=\textwidth]{figures/sections/math/temp_3f.png}
        \label{fig:fiber_example_cube}
    \end{subfigure}
    \label{fig:fiber_example}
    \caption{The fiber is expanded to include metadata fields that describe the semantics of $K$. In figure~\ref{fig:fiber_example_plane} the fiber is \textrm{temperature} $\times$ \textrm{time} and in figure~\ref{fig:fiber_example_cube} the fiber is \textrm{temperature}$\times$ \textrm{latitude} $\times$ {longitude}}.
\end{figure}

To encode the metadata, the fiber is expanded as illustrated in figure~\ref{fig:fiber_example}. The fiber in figure~\ref{fig:fiber_example_plane} is the cartesian product of the space of possible temperature values in degrees celsius and space of possible time values
\begin{equation}
F = [temp_{min}, temp_{max}] \times [time_{min}, time_{max}]
\label{eq:fiber_plane}
\end{equation}

while the fiber in figure~\ref{fig:fiber_example_cube} is the cartesian product of temperature, latitude, and longitude
\begin{equation}
F = [temp_{min}, temp_{max}] \times [-90, 90] \times [-180, 180]
\label{eg:fiber_cube}
\end{equation}

such that $E$ is the space of all possible points in $F$.

\begin{figure}[H]
    \includegraphics[width=1\linewidth]{figures/sections/math/fiberbundle.png}
    \label{fig:fiber_example_section}
    \caption{The section $\tau_1$ returns the blue points,  while $\tau_2$ returns the purple points. $\Gamma(E)$ is the set of all sections, including $\tau_1$ and $\tau_2$}.  
\end{figure}
Given the fiber described in equation~\ref{eq:fiber_plane}, the sections $\tau_{1}$ and $\tau_{2}$ in figure~\ref{fig:fiber_example_section} return tuples of the form
\begin{equation}
\tau(k) = (k, (temperature, time))
\end{equation}
such that sections with the constraint that time is monotonic return a timeseries.

@@

$M = M_{1} \times M_{2} \times \ldots \times M_{n}$
color X xpos (Cross product of monoid actions over V/E), cartesian product b/c independent. $M$ over $E$ was translated to acting over $\Gamma(V)$ of these act on $\Gamma(V)$
$\nu$ translates the sets, constraint on $\nu$ is equivariance of M. Is limited to the $Q(\Gamma(V))\subset of \Gamma(H)$ because not all visual tramsforms ($\mu$) are supported by all  $\rho$.

\begin{figure}
    \includegraphics[width=\textwidth]{figures/sections/math/diff_type_q.png}
\end{figure}
So we define the actions $M$ on the image of $Q$, $Q(\Gamma(V)) = Y$. We can backward define our actions?
